Verifying Spatial Properties of Stencil Computations

### 2016-11-16 paper.tex

\begin{definition}
  An \emph{index scheme} is a vector of size $n$ (an $n$-vector) with
  values drawn from $\domainVal = \mathbb{Z} \cup \{\infty\}$.  Values
  in $\mathbb{Z}$ represent the offsets of neighbourhood indices
  (\ie{}, the constant integer $c$ in an indexing expression of the
  form $i \pm c$, Def.~\ref{def:neighbour}) and $\infty$ is a wild-card
  representing both constant indices and unspecified indices. Throughout, $u, v,
  w$ range over
  index schemes and $u_i$ denotes the $i^{th}$ element of $u$. We
  write $(\domainVal)^n$ for the set of all $n$-dimensional schemes.
\end{definition}

Indexing schemes give an abstract
representation of array subscripts in the source language.
Our model represents the meaning of stencil specifications
as indexing schemes, which explain the range of possible
indexing behaviours which are consistent with the specification.
The relationship between source language
terms and indexing terms
 is given precisely by the following function \textsf{schematic},
 which forms part of checking and inference (\S\ref{sec:analysis}).

\newcommand{\schemeOne}{\textsf{s}_I}
%\newcommand{\schemeOne}{\textsf{schema$^1_I$}}
\newcommand{\schemeN}{\textsf{schematic}_I}
%\newcommand{\schemeN}{\textsf{schema$^n_I$}}
\begin{definition}%[Schematic]
\label{def:schematic}
Let $\schemeOne$ be a partial function, parameterised by a set of
induction variables $I$, mapping expression syntax to
$\domainVal$ values:
%%
\begin{align*}
\schemeOne(e)
 & =  \begin{cases}
c & e \equiv i + c \;\; \wedge \;\; i \in I \\
-c & e \equiv i - c \;\; \wedge \;\; i \in I \\
\infty & \textit{$I \not\subseteq \textit{FV}(e)$} \\
\bot   & \textit{otherwise}
\end{cases}
\end{align*}
where $\textit{FV}(e)$ is the set of free variables in $e$. In the
first two cases, expressions that are a constant offset of an
induction variable (\ie{}, neighbourhood indices) are mapped to the
offset $c$ or $-c$ $\in \domainVal$. In the third case, if the
expression is not a neighbourhood index, but neither is it a relative
index (\ie{}, it is not defined in terms of an induction variable) then it
is mapped to $\infty \in \domainVal{}$. Otherwise, $\schemeOne(e)$
is undefined (returns $\bot$).

@@

\begin{definition}[Generalisation]
Given two schemes $v, w$ of dimensionality $n$,
then $w$ is said to \emph{generalise} $v$
(or conversely $v$ \emph{specialises} $w$),
written $v \preceq_n w$, iff the following holds:
%%
\[
v \preceq_n w \;\; \Leftrightarrow  \;\;
  \forall i\!\in\!\{ 1, \ldots, n \} . \, (w_i = \infty \, \vee \, v_i = w_i)
\]
%%
That is, two schemes are related by the (reflexive and transitive) binary relation
$\preceq_n$ over $(\mathbb{Z}_{\infty})^n$ if in each dimension
the two schemes are equal or $w_i = \infty$
in place of some $\mathbb{Z}$ in $v_i$ for $v \preceq_n w$.
Thus, the element $\infty$ generalises any index in the same dimension.
%
\end{definition}

\begin{example}
Consider the following indexing schemes
corresponding to array indexing terms
$(i, j+1)$, $(1, j+1)$, $(3, j)$ and $(i*2, j)$:
%
\begin{align*}
\begin{array}{ll}
\schemeN{(i, j+1)} & =\vtwoh{0}{1} \\
\schemeN{(1, j+1)} & = \vtwoh{\infty}{1} \\
\schemeN{(3, j)} & = \vtwoh{\infty}{0} \\
\schemeN{(i*2, j)} & = \bot
\end{array}
\end{align*}
There is one generalisation between these
schemes, namely that $\vtwoh{0}{1} \preceq_2 \vtwoh{\infty}{1}$.
Note the last example is undefined as its first expression is a
non-neighbour relative index.
\end{example}

>>>>

\begin{definition}
  An \emph{index scheme} is a vector of size $n$ (an $n$-vector) with
  values drawn from $\domainVal = \mathbb{Z} \cup \{\infty\}$.  Values
  in $\mathbb{Z}$ represent the offsets of neighbourhood indices
  (\ie{}, the constant integer $c$ in an indexing expression of the
  form $i \pm c$, Def.~\ref{def:neighbour}) and $\infty$ is a wild-card
  representing both constant indices and unspecified indices.
  For an index scheme $u$, $u_i$ denotes the $i^{th}$ element of $u$. We
  write $(\domainVal)^n$ for the set of all $n$-dimensional schemes.
\end{definition}

Indexing schemes abstract array subscripts.
Our model represents the meaning of stencil specifications
as indexing schemes, which explain the range of possible
indexing behaviours which are consistent with the specification.
The relationship between source language
terms and indexing terms
 is given precisely by the following function \textsf{schematic},
 which forms part of checking and inference (\S\ref{sec:analysis}).

\newcommand{\schemeOne}{\textsf{s}_I}
%\newcommand{\schemeOne}{\textsf{schema$^1_I$}}
\newcommand{\schemeN}{\textsf{schematic}_I}
%\newcommand{\schemeN}{\textsf{schema$^n_I$}}
\begin{definition}%[Schematic]
\label{def:schematic}
Let $\schemeOne$ be a partial function, parameterised by a set of
induction variables $I$, mapping expression syntax to
$\domainVal$ values:
%%
\begin{align*}
\schemeOne(e)
 & =  \begin{cases}
c & e \equiv i + c \;\; \wedge \;\; i \in I \\
-c & e \equiv i - c \;\; \wedge \;\; i \in I \\
\infty & \textit{$I \cap \textit{FreeVariables}(e) = \emptyset $} \\
\bot   & \textit{otherwise}
\end{cases}
\end{align*}
In the
first two cases, expressions that are a constant offset of an
induction variable (\ie{}, neighbourhood indices) are mapped to the
offset $c$ or $-c$ $\in \domainVal$. In the third case, if the
expression is not defined in terms of an induction variable, it
is mapped to $\infty \in \domainVal{}$. Otherwise, $\schemeOne(e)$
is undefined (returns $\bot$).

@@

\begin{definition}[Generalisation]
Given two schemes $v, w$ of dimensionality $n$,
then $w$ is said to \emph{generalise} $v$
(or conversely $v$ \emph{specialises} $w$),
written $v \preceq_n w$, iff the following holds:
%%
\[
v \preceq_n w \;\; \Leftrightarrow  \;\;
  \forall i\!\in\!\{ 1, \ldots, n \} . \, (w_i = \infty \, \vee \, v_i = w_i)
\]
%%
Two schemes are related by the (reflexive and transitive) binary relation
$\preceq_n$ over $(\mathbb{Z}_{\infty})^n \times (\mathbb{Z}_{\infty})^n$.
The element $\infty$ generalises any index in the same dimension.
%
\end{definition}

\begin{example}
Consider the following array subscript to indexing schemes
conversions:
%
\begin{align*}
\begin{array}{ll}
\schemeN{(i, j+1)} & =\vtwoh{0}{1} \\
\schemeN{(1, j+1)} & = \vtwoh{\infty}{1} \\
\schemeN{(3, j)} & = \vtwoh{\infty}{0} \\
\schemeN{(i*2, j)} & = \bot
\end{array}
\end{align*}
There is one generalisation between these
schemes, namely that $\vtwoh{0}{1} \preceq_2 \vtwoh{\infty}{1}$.
The last example is undefined as its first expression is a
non-neighbour relative index.
\end{example}
